\section{Theory}\label{sec:theory}

We develop a theoretical framework grounded in stakeholder theory that incorporates political ideology as a critical dimension of stakeholder relationships. Our framework engages with three streams of literature: stakeholder theory and its assumptions about alignment, corporate political activity and stakeholder reactions, and research on political consumerism and employee ideology. We then develop predictions about how stakeholder ideological composition varies across firms and how alignment affects firm outcomes.

Our theoretical development speaks to what \citet{bridoux2023new} call the ``new stakeholder theory on organizational purpose,'' which grapples with questions of stakeholder enfranchisement and value distribution. By introducing ideological composition as a dimension of stakeholder relationships, we contribute to understanding how stakeholder preferences---and conflicts among them---shape organizational strategy.

\subsection{Stakeholder Theory and the Alignment Problem}

Stakeholder theory posits that firms should create value for all stakeholders, not just shareholders \citep{freeman1984strategic}. This perspective recognizes that employees, customers, suppliers, communities, and other groups affect and are affected by firm activities, and that successful organizations manage these relationships to generate mutual benefit. The instrumental version of stakeholder theory further argues that attending to stakeholder interests is not merely ethical but also strategically advantageous: firms that treat stakeholders well enjoy better long-term performance \citep{jones1995instrumental}.

A key insight from organizational theory is that managers occupy a central position in reconciling stakeholder claims. \citet{cyert1963behavioral} conceptualized organizations as coalitions of individuals with ``disparate demands, changing foci of attention, and limited ability to attend to all problems simultaneously.'' \citet{pfeffer1978external} extended this view, arguing that organizations are ``other-directed,'' influenced by actors who control critical resources and have managers' attention. \citet{hill1992stakeholder} synthesized these perspectives in their stakeholder-agency model, positioning managers at the center of a ``nexus of contracts'' among stakeholders. In their formulation, managers are unique: they alone enter contractual relationships with all other stakeholders and possess direct control over the firm's decision-making apparatus. Their central position carries responsibility for ``reconciling divergent interests by making strategic decisions and allocating strategic resources in a manner most consistent with the claims of stakeholder groups'' \citep[p.~134]{hill1992stakeholder}.

Central to this framework is the stakeholder salience model developed by \citet{mitchell1997toward}. Salience---the degree to which managers prioritize competing stakeholder claims---depends on three attributes: power (the stakeholder's ability to impose its will), legitimacy (the social appropriateness of the stakeholder's claim), and urgency (the time-sensitivity and criticality of the claim). Stakeholders possessing all three attributes demand immediate managerial attention; those with fewer attributes may be ignored or deprioritized. This model explains why managers attend to some stakeholders more than others, but it assumes that stakeholder preferences are relatively stable and knowable.

Recent work has advanced what \citet{mcgahan2021integrating} calls the ``new stakeholder theory'' (NST), which brings greater precision to understanding stakeholder involvement in organizations. The NST grapples with two canonical questions: which stakeholders are enfranchised in organizations, and how is value created through stakeholder collaboration distributed? \citet{bridoux2022stakeholder} trace how stakeholder theory has evolved from normative arguments about stakeholder consideration to formal analysis of how stakeholder composition affects strategic outcomes. A central insight is that stakeholders bind resources to organizations---employees contribute human capital, customers provide revenue, suppliers offer inputs---and understanding these resource contributions illuminates why stakeholder management is strategically important \citep{barney2018stakeholder}.

An implicit assumption in much stakeholder theory is that stakeholder interests, while diverse, can be reconciled or balanced. \citet{freeman2010stakeholder} argue that value creation need not be zero-sum: clever managers can find solutions that benefit multiple stakeholders simultaneously. This ``both/and'' perspective underpins stakeholder capitalism's normative appeal. Yet it sidesteps a difficult question: what happens when stakeholder interests are not merely different in magnitude but opposite in direction?

Political ideology presents precisely this challenge. When a firm's customers include both liberals and conservatives, satisfying one group's preferences may alienate the other. This is not a problem of incomplete information or misaligned incentives that can be solved through better contracts or communication. It is a fundamental conflict rooted in different values, worldviews, and group identities. The stakeholder framework, developed before polarization reached current levels, does not fully address this scenario.

Our research extends stakeholder theory by introducing ideological heterogeneity as a dimension of stakeholder composition. We examine not only whether firms have customers and employees but what those stakeholders believe politically---and whether different stakeholder groups share similar political orientations. This extension enables us to address the ``caught in the middle'' problem that firms increasingly face.

\subsection{Political Ideology as a Stakeholder Attribute}

Political ideology---an individual's system of beliefs about how society should be organized---shapes economic behavior in multiple ways. A large literature documents that partisan identity influences consumption patterns, employment choices, and reactions to firm activities. These effects operate through several mechanisms.

\textit{Value expression.} Consumers and employees derive utility from associating with organizations that share their values \citep{bhattacharya2003consumer}. Just as consumers may prefer products from firms with strong environmental commitments, they may prefer firms whose political positioning aligns with their own. The rise of ``political consumerism''---using market choices to express political preferences---reflects this tendency \citep{stolle2005consumers, endres2017resisting}. When consumers boycott firms over political stances or buycott firms they wish to support, they are expressing identity through market behavior.

\textit{In-group favoritism.} Social identity theory suggests that individuals favor members of their own groups and disfavor out-groups \citep{tajfel1979integrative}. When partisanship becomes a social identity---as it increasingly has in the United States \citep{mason2018one, iyengar2019origins}---consumers and employees may prefer organizations perceived as co-partisan. This extends beyond explicit political stances: even implicit cues about a firm's political orientation (its customer base, employee composition, geographic footprint) may trigger in-group preferences.

\textit{Trust and legitimacy.} Ideology shapes perceptions of organizational legitimacy. Conservative individuals may view firms with progressive workforces as less trustworthy; liberals may distrust firms associated with conservative causes. Because legitimacy affects willingness to transact---customers must believe the firm will fulfill its promises; employees must believe the firm will treat them fairly---ideological mismatch can erode the foundation of stakeholder relationships.

These mechanisms suggest that stakeholder political composition is not merely a demographic curiosity but a strategically relevant attribute. Firms serving politically homogeneous stakeholder bases may find it easier to maintain trust and satisfy expectations. Firms serving diverse bases face the challenge of appealing to groups with conflicting preferences.

Which mechanism dominates depends on context. Value expression is most salient when consumption or employment is publicly visible and identity-relevant---buying a Tesla signals environmental values; working at a nonprofit signals prosocial orientation. In-group favoritism intensifies when partisan identity is chronically accessible, as during elections or periods of high polarization. Trust and legitimacy concerns arise when stakeholders must make inferences about unobserved firm behavior---will this employer treat me fairly? Will this company honor its warranty? When political cues serve as heuristics for trustworthiness, even incidental signals of ideology can affect transactions.

\subsection{Sorting Equilibria and Stakeholder Selection}

The distribution of stakeholder ideology across firms is not random; it reflects equilibrium sorting processes. Just as labor markets exhibit assortative matching on skills and preferences, consumer and labor markets may exhibit assortative matching on ideology. Understanding these sorting dynamics is essential for interpreting observed patterns of stakeholder composition.

Consider consumer sorting. If liberals prefer firms perceived as progressive, and conservatives prefer firms perceived as traditional, then even small initial differences in firm positioning can be amplified through self-reinforcing dynamics. Early adopters attract like-minded customers, whose presence reinforces the firm's ideological image, attracting more similar customers while deterring dissimilar ones. The result is ideological clustering that exceeds what firm actions alone would produce.

Employee sorting operates through similar mechanisms but with additional frictions. Workers cannot costlessly switch employers; they face search costs, moving costs, and firm-specific human capital that create switching frictions. These frictions mean that employee ideological composition may be stickier than customer composition, adjusting more slowly to changes in firm positioning or local labor market conditions. Geographic constraints also matter: employees must live near their workplace (or accept remote arrangements), while customers can travel or shop online. This geographic tethering means employee ideology may more closely reflect local population ideology than customer ideology does.

The interaction between consumer and employee sorting creates complex equilibrium dynamics. A firm that attracts conservative customers may find it easier to recruit conservative employees (who prefer serving co-partisans), which further reinforces the firm's conservative image. Conversely, a firm with progressive employees may attract progressive customers who value interactions with like-minded workers. These reinforcing dynamics suggest that stakeholder alignment emerges endogenously through sorting, not merely from deliberate firm strategy.

This sorting perspective has important implications. First, observed stakeholder composition reflects both firm choices and stakeholder choices---it is jointly determined. Second, composition may exhibit path dependence: early positioning decisions have persistent effects through cumulative sorting. Third, composition changes face inertia: shifting stakeholder ideology requires overcoming the gravitational pull of existing stakeholders who attract similar others.

\subsection{Corporate Political Activity and Stakeholder Reactions}

A substantial literature examines corporate political activity (CPA)---the actions firms take to influence government policy---and its consequences \citep{hillman2004political}. CPA includes lobbying, political donations, direct engagement with policymakers, and public advocacy on political issues. Research has documented both benefits (favorable policy outcomes, regulatory access) and costs (stakeholder backlash, reputational damage) of political engagement \citep{hadani2017institutional}.

Recent work has increasingly focused on how stakeholders respond to corporate political involvement. \citet{burbano2021effect} finds that employees respond negatively to CEO political donations that conflict with their own preferences, with effects on job satisfaction and turnover intentions. \citet{hambrick2019political} documents that CEO political contributions affect customer perceptions and firm reputation. \citet{melloni2019corporate} shows that firms face pressure from employees and customers when they engage with political issues, and that responses depend on the political composition of these stakeholders.

One stream of this literature examines ``CEO activism''---public statements by corporate leaders on political and social issues \citep{chatterji2019origins}. While CEO activism can enhance brand loyalty among sympathetic customers and increase employee engagement among like-minded workers, it risks alienating stakeholders on the other side. \citet{korschun2022taking} theorizes that CEO activism represents a risky bet: the firm gains among stakeholders who share the CEO's position but loses among those who disagree.

A deeper question is whether CEO activism can be profitable despite this trade-off. \citet{melloni2023cashing} develop a formal model showing that credible, value-enhancing CEO activism is possible, but only under specific conditions. Their key insight is that controversy is a \textit{feature}, not a bug, of effective activism. Consumers understand that CEOs may engage in ``wokewashing''---making false or misleading statements to pander to valuable demographics. For activism to credibly signal firm type, it must be costly enough to deter imitation by firms that do not genuinely share those values. This cost comes from alienating stakeholders on the opposing side. Paradoxically, activism that generates no backlash signals nothing, because it could be costlessly adopted by any firm. Only activism that sacrifices some stakeholders can credibly appeal to others.

This logic extends to competitive dynamics. \citet{mohliver2023corporate} introduce the concept of ``corporate social counterpositioning,'' in which firms deliberately take opposing stances on polarizing social issues to differentiate from competitors. Unlike conventional wisdom that portrays CSR as universally positive, their theory recognizes that highly salient, polarizing issues create opportunities for horizontal differentiation. When a rival adopts a progressive position on gun control or LGBTQ rights, a firm may strategically counterposition by opposing that stance---not despite, but because of, the controversy it generates. The profitability of counterpositioning depends on issue attributes (salience and polarization) and market structure. In fragmented markets where firms compete for distinct customer segments, counterpositioning can be an effective differentiation strategy. This framework helps explain why some firms embrace politically divisive positions rather than pursuing broad, inoffensive CSR.

This research highlights a critical point: the consequences of corporate political engagement depend on stakeholder composition. A firm with a homogeneously liberal customer base may benefit from progressive CEO activism; the same activism could devastate a firm serving conservatives. Recent work by \citet{mckean2024ideologies} examines this directly, finding that ideological alignment between a firm's upper echelons and its stakeholders influences participation in progressive corporate activism. Similarly, \citet{gupta2025political} theorize how growing political polarization reshapes firms' nonmarket strategies over the policy life cycle, showing that polarization delays social consensus and intensifies both activism and industry resistance. Yet existing research typically treats stakeholder composition as unobserved, inferring it indirectly from reactions or modeling it theoretically. We contribute by directly measuring stakeholder ideology, enabling precise analysis of how composition moderates responses to political activity.

\subsection{The Employee Dimension}

While much research on stakeholder politics focuses on consumers, employees represent an equally important stakeholder group. Employees' political beliefs shape their workplace experiences, organizational commitment, and willingness to engage in discretionary effort.

Value congruence---the alignment between individual and organizational values---is a well-established predictor of employee outcomes \citep{kristof1996person, edwards2009person}. When employees perceive that their employer shares their values, they report higher job satisfaction, stronger organizational identification, and lower turnover intentions. Extending this to political values, we would expect employees whose political orientation aligns with their employer's (or their coworkers') to experience greater fit and commitment.

Recent work has begun to examine political dynamics in the workplace directly. \citet{sisco2022managing} finds that political diversity in workplaces can generate conflict, with effects on team performance and employee well-being. \citet{burbano2021effect} documents that employees care about their employers' political activities and may exit when those activities conflict with their beliefs. The Politics at Work project \citep{mcconnell2022economic} reveals substantial variation in employee partisan composition across firms and shows that this composition predicts corporate political behavior.

An important question is whether employee and customer political composition covary. If they do, firms may face reinforcing pressures from aligned stakeholder groups. If they diverge, firms must navigate cross-stakeholder tensions---progressive employees serving conservative customers, or vice versa. This divergence could create internal conflict, affect customer service, and complicate corporate political positioning. We provide the first large-scale evidence on the correlation between employee and customer ideology.

\subsection{Within-Stakeholder Diversity}

Beyond cross-stakeholder alignment, we consider within-stakeholder diversity. A firm's customers are not a homogeneous group; they may themselves hold diverse political views. Similarly, a firm's employees may span the political spectrum. The degree of within-group heterogeneity has implications for stakeholder management.

\citet{harrison2007whats} distinguish three types of diversity: separation (differences in position along a continuum), variety (differences in kind or category), and disparity (differences in concentration of valued resources). Political ideology represents separation diversity---stakeholders differ in their positions along the liberal-conservative spectrum. This distinction matters because separation diversity, unlike variety, tends to generate conflict rather than complementarity. When group members hold opposing positions on valued issues, interaction becomes fraught; compromise positions satisfy no one.

Consider two firms with the same mean customer partisan lean of 50\% Republican. Firm A's customers are uniformly distributed: half are strong Republicans, half are strong Democrats. Firm B's customers cluster near the center: most are moderates or weak partisans. Though both firms have the same average, their strategic situations differ dramatically. Firm A faces the ``caught in the middle'' problem acutely: any political move will alienate half its customers. Firm B, serving a politically moderate base, may have more latitude.

This insight connects to research on market positioning under heterogeneous preferences \citep{caves1977industrial}. Firms may differentiate to serve distinct segments, pursue broad appeal strategies, or attempt niche positioning. When heterogeneity is political, differentiation becomes more fraught: unlike product preferences, political preferences are tied to identity and group membership. Attempting to serve ``both sides'' may be perceived as inauthentic or unprincipled, pleasing neither group.

We therefore examine not only mean stakeholder ideology but its dispersion. Firms with more politically diverse customer bases face different challenges than those serving homogeneous markets, and this diversity may moderate the relationship between ideological alignment and performance.

\subsection{Temporal Variation: The Role of Political Climate}

The salience of political ideology as a stakeholder attribute varies over time. During periods of elevated political conflict---contested elections, Supreme Court decisions, highly publicized political events---partisan identity becomes more accessible and influential. During quieter periods, other considerations may dominate consumer and employee behavior.

This temporal variation provides an identification opportunity. If stakeholder ideological composition affects firm performance, we would expect this effect to be stronger when political issues are salient. When the ``partisan temperature'' is high, consumers are more likely to factor politics into purchasing decisions, employees are more likely to evaluate employers through a political lens, and misalignment should matter more. When the temperature is low, ideology recedes in importance, and alignment or misalignment should have weaker effects.

We exploit this intuition using the Partisan Conflict Index (PCI), a monthly measure of political disagreement developed by the Federal Reserve Bank of Philadelphia \citep{azzimonti2018partisan}. The PCI captures the intensity of partisan conflict based on news coverage of political disagreements. By interacting stakeholder ideological alignment with the PCI, we can test whether alignment matters more during periods of heightened political salience.

This approach addresses endogeneity concerns that plague cross-sectional comparisons. One might worry that firms with more aligned stakeholders differ from misaligned firms in unobservable ways that also affect performance. But if the alignment-performance relationship varies with political climate---stronger when PCI is high, weaker when PCI is low---this suggests the relationship is driven by political mechanisms rather than confounds. The interaction provides quasi-experimental leverage on a fundamentally correlational question.

\subsection{Scope Condition: The Information Environment}

Our theoretical mechanisms require that stakeholders have some awareness of firm ideology or the ideology of fellow stakeholders. This information environment represents an important scope condition. When firm ideology is opaque---unknown to customers, invisible to potential employees---the mechanisms we describe cannot operate. Stakeholders cannot sort on ideology they do not perceive; they cannot reward alignment they do not recognize.

Several factors affect ideological visibility. First, firm size and prominence matter: large public companies receive more media scrutiny, making their political activities and workforce composition more salient. Second, industry context matters: firms in politically charged sectors (energy, firearms, media) face greater attention to their political positioning than firms in neutral sectors. Third, CEO visibility affects transparency: activist CEOs who speak publicly about political issues make firm ideology more observable than CEOs who remain silent.

The rise of social media and online reviews may be increasing ideological transparency across the board. Consumers can now easily learn about firms' political donations, employee reviews can reveal workplace political climate, and viral moments can rapidly publicize a firm's perceived ideology. This increased transparency may be strengthening the mechanisms we describe, making stakeholder ideology more consequential than it was in earlier, less informationally connected eras.

We do not directly test how the information environment moderates our results, but we acknowledge this as a boundary condition. Our findings likely understate the potential effects of ideology in settings with high transparency and overstate them in settings where firm ideology remains hidden from stakeholders.

\subsection{Summary of Theoretical Predictions}

Our theoretical framework generates several predictions, which we organize by the type of analysis required to test them.

First, we expect substantial variation in stakeholder ideological composition across firms. While geography will explain much of this variation---businesses in Republican areas attract Republican visitors---we predict that industry, brand, and business type will also matter. Certain categories (e.g., organic grocers, gun shops) should attract politically distinct customer bases even within the same geographic area.

Second, we expect employee and consumer political composition to be positively correlated but imperfectly aligned. Firms in conservative areas will tend to have both conservative employees and conservative customers, generating positive correlation. But the correlation should be imperfect because employees and customers are selected through different processes: employees choose employers based on job availability, compensation, and career opportunities; customers choose businesses based on product offerings, convenience, and price. These distinct selection processes should introduce divergence.

Third, we expect stakeholder ideological alignment to be associated with firm performance, particularly during periods of elevated political conflict. When employees and customers share political orientations, the firm faces less tension in its stakeholder management and can more easily maintain trust across groups. When they diverge, the firm must navigate conflicting expectations, risking backlash from one group or the other. This relationship should intensify when political issues are salient, as captured by high PCI.

Fourth, we expect within-stakeholder diversity to moderate the alignment-performance relationship, with higher diversity \textit{attenuating} the benefits of alignment. The mechanism is aggregation: when a firm's customers span the political spectrum, positive reactions from aligned customers are offset by negative reactions from misaligned customers, dampening the net effect. Conversely, firms with ideologically homogeneous customer bases face more concentrated reactions---strongly positive when aligned, strongly negative when misaligned. We therefore predict that the alignment-performance relationship is steeper (more positive slope) for firms with low customer ideological diversity than for firms with high diversity.

These predictions guide our empirical analysis. We turn next to describing the data and methods used to test them.
